A Proposal for Multi-category Lexical and Sentential Similarity Datasets Based on Commonality Lists

### 2016-06-14 paper.tex
New title and edits to the section on sentences.

\title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}

@@

% \section{Similarity in Context}

% In the previous section we focused on the properties of similarity, illustrating it with word-word examples. This section is dedicated to the nuances of similarity measurement between words, phrases and sentences. Here, we assume that similarity requires the agreement of the grammatical roles.

% \section{Constructions and Language Sub-systems}
\section{Sentence Similarity}

@@

To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them. So what are they?

Words, sentences and other constructions draw attention to states of affairs around us. This is both pragmatically and semantically speaking the case. As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.

The point here is that some sentences are similar to others with respect to the \emph{functions} they perform. For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.

% As pointed out above, once one has categories one can speak of similarity.

Thus, the amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence. \newcite{tomasello1998return} points out that sentence-level constructions show prototype effects similar to those discussed above for lexical systems (e.g.~colours). Consider the following sentences:
\begin{compactitem}
    \item \textit{``John gave Mary the book.''} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
      \item \textit{``John promised Mary the book.''} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
    %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
    %   \item \textit{``John left Mary the book''} is a example of an \emph{Future Transfer} construction. These usually are build around words such as \textit{bequeath, leave, reserve, grant, etc.}
    %   \item \textit{``John allowed Mary one book''} is a example of an \emph{Enabling Conditions for Transfer} construction. These usually are build around words such as \textit{permit, allow, etc.}
    %   \item \textit{``John wrote Mary a book''} is a example of an \emph{Intended Transfer} construction. These usually are build around words such as \textit{bake, build, make, get, grab, win, earn, etc.}
\end{compactitem}

@@

However, these categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}. The same similarity principle can be applied to phrases as to words. In this case, similarity is measured in context, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}. With verbs this idea can be applied to compare transitive verbs with intransitive. For example, \textit{to cycle} is similar to \textit{to ride a bicycle}.

@@

But still, what about pragmatics? As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world. For example, he points to the fact that if a person asks another person to \textit{``pass the wine''} they are actually asking for the \textit{bottle}. The speaker just attracts attention to an object of perception in a given situation.

@@

@@

@@

@@

In this contribution we discussed the notion of ``similarity'' from an interdisciplinary perspective. We contrasted properties of the similarity relation described in the field of psychology with the characteristics of similarity datasets used in computational linguistics. This lead to the recommendations on how to improve the later by removing low score ambiguity in a multi-category similarity dataset.

>>>>

% \title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}
\title{A Proposal for Multi-category Lexical and Sentential Similarity Datasets Based on Commonality Lists}

@@

\section{Sentence Similarity}

@@

To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them. So what are they? First of all, words, sentences and other constructions draw attention to \emph{states of affairs} around us. 
% This is both pragmatically and semantically speaking the case.

\paragraph{Prototype effect}

Some sentences are similar to others with respect to the functions they perform. As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{X floosed Y the Z} and \textit{X was floosed by Y}. This is due to their similarity to sentences such as \textit{John gave Mary the book} and \textit{Mary was kissed by John} respectively. Thus, \textit{X floosed Y the Z} is clearly a \emph{transfer of possession} or \emph{dative} \cite{bresnan2007predicting}.

The amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence. \newcite{tomasello1998return} points out that sentence-level constructions show \emph{prototype effects} similar to those discussed above for lexical systems (e.g.~colours). Consider the following sentences:
\begin{compactitem}
    \item \textit{John gave Mary the book.} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
      \item \textit{John promised Mary the book.} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
    %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
    %   \item \textit{``John left Mary the book''} is a example of an \emph{Future Transfer} construction. These usually are build around words such as \textit{bequeath, leave, reserve, grant, etc.}
    %   \item \textit{``John allowed Mary one book''} is a example of an \emph{Enabling Conditions for Transfer} construction. These usually are build around words such as \textit{permit, allow, etc.}
    %   \item \textit{``John wrote Mary a book''} is a example of an \emph{Intended Transfer} construction. These usually are build around words such as \textit{bake, build, make, get, grab, win, earn, etc.}
\end{compactitem}

@@

\paragraph{Similarity in Context}

However, prototype categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}. The same similarity principle can be applied to phrases as to words. In this case, similarity is measured in context, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}. With verbs this idea can be applied to compare transitive verbs with intransitive. For example, \textit{to cycle} is similar to \textit{to ride a bicycle}.

@@

\paragraph{Attention Attraction}

But still, what about pragmatics? As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world. For example, he points to the fact that if a person asks another person to \textit{pass the wine} they are actually asking for the \textit{bottle}. The speaker just attracts attention to an object of perception in a given situation.

\paragraph{Grammaticalisaton and Lexicalisaton}

@@

\paragraph{Discourse Fit}

@@

\paragraph{Language Sub-systems and Strategies}

@@

Above we identified seven criteria by which sentence similarity can be compared.

@@

In this contribution we discussed the notion of similarity from an interdisciplinary perspective. We contrasted properties of the similarity relation described in the field of psychology with the characteristics of similarity datasets used in computational linguistics. This lead to the recommendations on how to improve the later by removing low score ambiguity in a multi-category similarity dataset.
